% !Mode:: "TeX:DE:UTF-8:Main"
%
%
%JOURNAL CODE  SEE DOCUMENTATION
\documentclass[biber,draft]{nowfnt} % creates the journal version, needs biber version
%wrapper for book and ebook are created automatically.
\usepackage{imakeidx}
\usepackage[utf8]{inputenc}

\def\withcolors{1}
\def\withnotes{1}
\usepackage{ccanonne}

\usepackage{framed}
\usepackage{tabularx}

%%%%%%%%%%%%%%%%% Exercises %%%%%%%%%%%%%%%%%
\newtheorem{question}{Exercise}[chapter]

\newcommand{\exinline}[1]{(\refstepcounter{question}Exercise~\thequestion\label{#1})}

\crefname{question}{Exercise}{Exercises}
\Crefname{question}{Exercise}{Exercises}
\crefname{claim}{Claim}{Claims}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\newcolumntype{Y}{>{\centering\arraybackslash}X}
\renewcommand\tabularxcolumn[1]{m{#1}}% for vertical centering text in X column


\usepackage{physics} %% <- redefines var
\usepackage[fulladjust]{marginnote}
\renewcommand*{\marginfont}{\scriptsize}
\setlength{\marginparwidth}{18mm}
\newcommand{\exercise}[1]{\marginnote{\textbf{E:} #1}}

\newcommand{\tbc}{\noindent\hl{\sc{}to be continued}\xspace}


\newcommand{\nss}{\textcolor{red}{\ns_1}}
\newcommand{\nst}{\textcolor{red}{\ns_2}}
\newcommand{\occur}{\textcolor{red}{N}}

%ARTICLE TITLE
\title{Distribution Testing, From Stats to CS (and Back?)}


%ARTICLE SUB-TITLE
\subtitle{Just a working title}


%AUTHORS FOR COVER PAGE 
% separate authors by \and, item by \\
% Don't use verbatim or problematic symbols.
% _ in mail address should be entered as \_
% Pay attention to large mail-addresses ...

%if there are many author twocolumn mode can be activated.
%\booltrue{authortwocolumn} %SEE DOCUMENTATION
\maintitleauthorlist{
Cl\'ement L. Canonne \\
University of Sydney\\
clement.canonne@sydney.edu.au
}

%ISSUE DATA AS PROVIDED BY NOW
\issuesetup
{%
 copyrightowner={A.~Heezemans and M.~Casey},
 volume        = xx,
 issue         = xx,
 pubyear       = 2018,
 isbn          = xxx-x-xxxxx-xxx-x,
 eisbn         = xxx-x-xxxxx-xxx-x,
 doi           = 10.1561/XXXXXXXXX,
 firstpage     = 1, %Explain
 lastpage      = 18
 }

%BIBLIOGRAPHY FILE
%\addbibresource{sample.bib}
%\addbibresource{testfnt.bib}
\addbibresource{bibliography.bib}

% \usepackage{mwe}

%AUTHORS FOR ABSTRACT PAGE
\author[1]{Cl\'ement L. Canonne}

\affil[1]{University of Sydney; clement.canonne@sydney.edu.au}

\articledatabox{\nowfntstandardcitation}

\begin{document}
\makeabstracttitle
\begin{abstract}
We focus on some specific problems in distribution testing,
taking goodness-of-fit as a running example. In particular,
we do not aim to provide a comprehensive summary of all
the topics in the area; but will provide self-
contained proofs and derivations of the main results, trying
to highlight the unifying techniques.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{What is distribution testing?}
\input{chap-what}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Testing goodness-of-fit of univariate distributions}
  \label{chap:identity}

\cnote{In-depth chapter with derivation of many results. Focus on one-sample
testing.}
\input{chap-identity}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{General algorithmic frameworks}
\section{Testing-by-learning}
\section{Shape restrictions}
\section{Fourier approach}
\section{The $\lp[2]$ testing framework of Diakonikolas--Kane}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Testing high-dimensional distributions}

\section{Testing product Bernoulli and Gaussian distributions}
\cnote{Mention the hardness of robust testing (Diakonikolas--Kane--Stewart),
to show the importance of the assumption.}

\section{Bayesian networks}
\section{Ising models and Markov Random Fields}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Testing with Constrained Measurements}
\cnote{Discuss what those can be (local privacy, noisy channels, linear of
partial measurements, etc). Focus on communication constraints ($\numbits$-bit
measurements). Talk about adaptive schemes?}
\section{Continuous distributions and quantization}
\section{Random hashing and domain compression}
\section{Application: distributed inference}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Information-theoretic lower bounds}
  \label{chap:lowerbounds}
\section{Proving hardness by reductions}
\section{Indistinguishability via moment-matching}
\section{Ingster's method}
\section{Application: distributed inference}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Beyond \iid sampling}
\cnote{Less detailed: This is a short chapter, where each section contains one key result/example
to illustrate the model or question, along with references.}

\section{Conditional sampling}
\cnote{Ability to sample, conditioning on specific events.}

\section{Evaluation access}
\cnote{Exact or approximate access to the probability mass or cumulative
distribution function.}

\section{Testing properties of Markov chains}
\cnote{Only one sample (trajectory) of a Markov chain.}

\section{Causal models and interventions}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Quantum distribution testing}

\cnote{Less detailed: Not many full derivations here, mostly statement of the results and
sketch of the proofs/ideas. Focus on testing the fully mixed state.}
\section{What is quantum distribution testing?}
\cnote{Brief (very) intro to quantum computing; mention of different models
(treated below), and their respective advantages/disadvantages.}
\section{Testing classical distributions with quantum measurements}
\cnote{Short section. This is mostly about getting faster algorithms, discuss
the (lack of) implications for sample complexity.}
\section{Testing quantum states\dots}
\subsection{With local measurements}
\subsection{With unentangled measurements}
\cnote{Two flavors: adaptive and non-adaptive. Discuss the recent work of
Bubeck, Chen, and Li showing a separation between the two.}
\subsection{With entangled measurements}
\cnote{Tight bounds for testing mixed states: in particular, O'Donnell and
Wright.}

\appendix

%BACKMATTER SEE DOCUMENTATION
\backmatter  % references, restarts sample

\printbibliography
% \printindex

\end{document}
